Numerical simulations of the Cordilleran ice sheet through the last glacial cycle

### 2015-05-28 cordillera-cycle.tex
Add location tag for the Omineca Mountains.

The possibility of late glacial readvances in the North American Cordillera has
been debated for some time \citep{Osborn.Gerloff.1997}, and locally these have
been reconstructed and dated. Radiocarbon-dated end moraines in the Fraser and
Squamish valleys, off the southern tip of the Coast Mountains, indicate
consecutive glacier maxima, or standstills while in overall retreat, one of
which corresponds to the Younger Dryas chronozone \citep{Clague.etal.1997,
Friele.Clague.2002, Friele.Clague.2002a, Kovanen.2002,
Kovanen.Easterbrook.2002}. Although most of these moraines characterise
independent valley glaciers, that were probably disconnected from the waning
Cordilleran ice sheet, the Finlay River
area of the northern Rocky Mountains presents a different kind of evidence.
There, sharp-crested moraines indicate a late-glacial readvance of local alpine
glaciers and, more importantly, their interaction with larger, lingering
remnants of the main body of the Cordilleran ice sheet in the valleys
\citep{Lakeman.etal.2008}. Additional evidence for late-glacial
alpine glacier readvances includes moraines in the eastern Coast Mountains,
Rocky Mountains and the Columbia Mountains \citep{Osborn.Gerloff.1997,
Menounos.etal.2008}.

Although further work is needed to constrain the timing of the late-glacial
readvance, to assess its extents and geographical distribution, and to identify
the potential climatic triggers \citep{Menounos.etal.2008}, it is interesting
to note that the simulation driven by the GRIP record produces a late-glacial
readvance in the Coast Mountains, Rocky Mountains and the Finlay River area,
corresponding to where it has been identified in the geological record
(Fig.~\ref{fig:deglac}, left panel). In addition to matching the location of
local readvances, the GRIP-driven simulation shows a large remnant of the
decaying ice sheet, centred on the Skeena Mountains, may still have existed
at the time of this late-glacial readvance.
In contrast, the EPICA-driven simulation produces a nearly-continuous
deglaciation with only a tightly restricted late-glacial readvance on the
western slopes of the Saint-Elias and Coast mountains
(Fig.~\ref{fig:deglac}, right panel).

>>>>

The possibility of late glacial readvances in the North American Cordillera has
been debated for some time \citep{Osborn.Gerloff.1997}, and locally these have
been reconstructed and dated. Radiocarbon-dated end moraines in the Fraser and
Squamish valleys, off the southern tip of the Coast Mountains, indicate
consecutive glacier maxima, or standstills while in overall retreat, one of
which corresponds to the Younger Dryas chronozone \citep{Clague.etal.1997,
Friele.Clague.2002, Friele.Clague.2002a, Kovanen.2002,
Kovanen.Easterbrook.2002}. Although most of these moraines characterise
independent valley glaciers, that may have been disconnected from the waning
Cordilleran ice sheet, the Finlay River area in the Omenica Mountains
(Fig.~\ref{fig:deglac}, OM) presents a different kind of evidence.
There, sharp-crested moraines indicate a late-glacial readvance of local alpine
glaciers and, more importantly, their interaction with larger, lingering
remnants of the main body of the Cordilleran ice sheet in the valleys
\citep{Lakeman.etal.2008}. Additional evidence for late-glacial
alpine glacier readvances includes moraines in the eastern Coast Mountains,
Rocky Mountains and the Columbia Mountains \citep{Osborn.Gerloff.1997,
Menounos.etal.2008}.

Although further work is needed to constrain the timing of the late-glacial
readvance, to assess its extents and geographical distribution, and to identify
the potential climatic triggers \citep{Menounos.etal.2008}, it is interesting
to note that the simulation driven by the GRIP record produces a late-glacial
readvance in the Coast Mountains and Rocky Mountains (Fig.~\ref{fig:deglac},
left panel). In addition to matching the location of some local readvances, the
GRIP-driven simulation shows that a large remnant of the decaying ice sheet
may still have existed at the time of this late-glacial readvance.
In contrast, the EPICA-driven simulation produces a nearly-continuous
deglaciation with only a tightly restricted late-glacial readvance on the
western slopes of the Saint-Elias and Coast mountains
(Fig.~\ref{fig:deglac}, right panel).
